Computable Poincar\'e--Friedrichs constants for the $L^{p}$ de~Rham complex over convex domains and domains with shellable triangulations

### 2025-01-31 main.tex
recursive estimate

We use that formalism to describe an estimate for the Poincar\'e--Friedrichs constant of the gradient potential, unwrapping the recursion.

@@

    Suppose that $n$-simplices are enumerated as $T_0, T_1, \dots, T_M$,
    and that with each $n$-simplex $T_{m} \in \calT$ 
    we have a sequence of indices 
    \begin{gather*}
        0 = \imath_{m,0}, \; \imath_{m,1}, \; \dots, \; \imath_{m,L_m} = m
    \end{gather*}
    such that for each $0 \leq m \leq M$ we have estimates of the following form:
    \begin{align*}
        \| w \|_{L^{p}(T_{0})} &\leq A_{m,0} \| \nabla u \|_{L^{p}(T_{0})}
        ,
        \\
        \| w \|_{L^{p}(T_{m,\ell})}
        &\leq  
        A_{m,\ell}
        \| w \|_{L^{p}(T_{m,\ell-1})} 
        +
        B'_{m,\ell}
        \| \nabla u \|_{L^{p}(T_{m,\ell  })} 
        + 
        B''_{m,\ell}
        \| \nabla u \|_{L^{p}(T_{m,\ell-1})}, 
        \qquad 1 \leq \ell \leq L_m
        .
    \end{align*}
    % TODO: complete here. We need a spanning tree. 
\end{theorem}
\begin{proof}
    Unwrapping the recursion for the norm over the $m$-th simplex, we get a total expression of the form 
    \begin{align}\label{math:fullrecursivesum}
        \begin{split}
            \| w \|_{L^{p}(T_{m})} 
            &
            \leq 
            \sum_{\ell=0}^{m}   \left( B_{m,m} \cdots B_{m,\ell+1} \right) A_{m,\ell} \| \nabla u \|_{L^{p}(T_{m,\ell})}
            +
            \sum_{\ell=0}^{m-1} \left( B_{m,m} \cdots B_{m,\ell+2} \right) A'_{m,\ell+1} \| \nabla u \|_{L^{p}(T_{m,\ell})}
            \\&
            = 
            A_{m,m} \| \nabla u \|_{L^{p}(T_{m,m})}
            +
            \sum_{\ell=0}^{m-1} \left( B_{m,m} \cdots B_{m,\ell+2} \right) \left( B_{m,\ell+1} A_{m,\ell} + A'_{m,\ell+1} \right) \| \nabla u \|_{L^{p}(T_{m,\ell})}
            \\&
            =:
            \sum_{\ell=0}^{M} C_{m,\ell} \| \nabla u \|_{L^{p}(T_{\ell})}
            .
        \end{split}
    \end{align}
    Here, $C_{m,\ell}$ is the coefficient of $\| \nabla u \|_{L^{p}(T_{\ell})}$, possibly zero, as it appears in the unwrapped recursive estimate of $\| w \|_{L^{p}(T_{m})}$. 
    By definition, $C_{m,0} = A_0$, and 
    \begin{align}
        C_{m,}
    \end{align}

    The global Poincar\'e--Friedrichs inequality follows via H\"older's inequality:
    \begin{align*}
        \begin{split}
            \| w \|_{L^{p}(\Omega)}^{p}
            &
            \leq 
            \sum_{m=0}^{M}
            \| w \|_{L^{p}(T_{m})}^{p}
            \\&
            \leq 
            \sum_{m=0}^{M}
            \left( \sum_{\ell=0}^{M} C_{m,\ell} \| \nabla u \|_{L^{p}(T_{\ell})} \right)^{p}
            \\&
            \leq 
            \sum_{m=0}^{M}
            \left( \sum_{\ell'=0}^{M} C_{m,\ell'}^{q} \right)^{\frac p q}
            \sum_{\ell=0}^{M} 
            \| \nabla u \|_{L^{p}(T_{\ell})}^{p} 
            \leq 
            \left(
                \sum_{m=0}^{M}
                \left( \sum_{\ell=0}^{M} C_{m,\ell}^{q} \right)^{\frac p q}
            \right)
            \| \nabla u \|_{L^{p}(\Omega)}^{p} 
            ,
        \end{split}
    \end{align*}
    where $q \in [1,\infty]$ satisfies $1 = 1/p + 1/q$ and with obvious modifications if $p=1$ or $p=\infty$. 
\end{proof}

>>>>

We use that formalism to describe an estimate for the Poincar\'e--Friedrichs constant of the gradient potential, unwrapping the recursion.
While the notation is a bit technical, the underlying idea is this:
Unrolling the recursion gives estimates for the potential over a simplex in terms of the potential over previous simplices. 
The final constant can be estimated by a norm of the vector of the coefficients that appear in the recursion.

@@

    Suppose that $n$-simplices are enumerated as $T_0, T_1, \dots, T_M$,
    and that with each $n$-simplex $T_{m} \in \calT$ 
    we have a sequence of indices 
    \begin{gather*}
        0 = \imath(m,0), \; \imath(m,1), \; \dots, \; \imath(m,L_m) = m
    \end{gather*}
    such that for each $0 \leq m \leq M$ we have estimates of the following form:
    \begin{align*}
        \| w \|_{L^{p}(T_{0})} &\leq A_{m,0} \| \nabla u \|_{L^{p}(T_{0})}
        ,
    \end{align*}
    \begin{align*}
        \| w \|_{ L^{p}(T_{\imath(m,\ell)}) }
        &\leq  
        A_{m,\imath(m,\ell)}
        \| w \|_{ L^{p}(T_{\imath(m,\ell-1)}) }
        \\&\qquad 
        +
        B'_{m,\imath(m,\ell)}
        \| \nabla u \|_{ L^{p}(T_{\imath(m,\ell)}) }
        + 
        B''_{m,\imath(m,\ell)}
        \| \nabla u \|_{ L^{p}(T_{\imath(m,\ell-1)}) }
        ,
        \quad 1 \leq \ell \leq L_m
        .
    \end{align*}
    Then 
    \begin{gather*}
        \| w \|_{ L^{p}(\Omega) }
        \leq 
        \left(
            \sum_{m=0}^{M}
            \left( \sum_{\ell=0}^{L_m} D_{m,\ell}^{q} \right)^{\frac p q}
        \right)^{\frac 1 p}
        \| \nabla u \|_{ L^{p}(\Omega) }
        ,
    \end{gather*}
    where 
    \begin{gather*}
        D_{m,L_m} = A_{m,m},
        \\
        D_{m,\ell} 
        = 
        \left( B_{m,\imath(m,L_m)} \cdots B_{m,\imath(m,\ell+2)} \right) 
        \left( B_{m,\imath(m,\ell+1)} A_{m,\imath(m,\ell)} + A'_{m,\imath(m,\ell+1)} \right) 
        ,
        \qquad 
        1 \leq \ell \leq L_{m}-1
        .
    \end{align*}
    % TODO: complete here. We need a spanning tree. 
\end{theorem}
\begin{proof}
    Unwrapping the recursion for the norm over the $m$-th simplex, we get a total expression of the form 
    \begin{align*}%\label{math:fullrecursivesum}
        \begin{split}
            \| w \|_{L^{p}(T_{m})} 
            &
            \leq 
            \sum_{\ell=0}^{L_m}   \left( B_{m,\imath(m,L_m)} \cdots B_{m,\imath(m,\ell+1)} \right) A_ {m,\imath(m,\ell  )} \| \nabla u \|_{L^{p}(T_{\imath(m,\ell)})}
            \\&\qquad 
            +
            \sum_{\ell=0}^{L_m-1} \left( B_{m,\imath(m,L_m)} \cdots B_{m,\imath(m,\ell+2)} \right) A'_{m,\imath(m,\ell+1)} \| \nabla u \|_{L^{p}(T_{\imath(m,\ell)})}
            \\&
            = 
            A_{m,m} \| \nabla u \|_{L^{p}(T_{\imath(m,L_m)})}
            \\&\qquad 
            +
            \sum_{\ell=0}^{L_m-1} 
            \left( B_{m,\imath(m,L_m)} \cdots B_{m,\imath(m,\ell+2)} \right) 
            \left( B_{m,\imath(m,\ell+1)} A_{m,\imath(m,\ell)} + A'_{m,\imath(m,\ell+1)} \right) 
            \| \nabla u \|_{L^{p}(T_{\imath(m,\ell)})}
            \\&
            =:
            \sum_{\ell=0}^{M} C_{m,\ell} \| \nabla u \|_{L^{p}(T_{\ell})}
            .
        \end{split}
    \end{align*}
    Here, $C_{m,\ell}$ is the coefficient of $\| \nabla u \|_{L^{p}(T_{\ell})}$, possibly zero, as it appears in the unwrapped recursive estimate of $\| w \|_{L^{p}(T_{m})}$. 
    The global Poincar\'e--Friedrichs inequality follows via H\"older's inequality:
    \begin{align*}
        \begin{split}
            \| w \|_{L^{p}(\Omega)}^{p}
            &
            \leq 
            \sum_{m=0}^{M}
            \| w \|_{L^{p}(T_{m})}^{p}
            \\&
            \leq 
            \sum_{m=0}^{M}
            \left( \sum_{\ell=0}^{M} C_{m,\ell} \| \nabla u \|_{L^{p}(T_{\ell})} \right)^{p}
            \\&
            \leq 
            \sum_{m=0}^{M}
            \left( \sum_{\ell=0}^{M} C_{m,\ell}^{q} \right)^{\frac p q}
            \sum_{\ell'=0}^{M} \| \nabla u \|_{L^{p}(T_{\ell'})}^{p} 
            \leq 
            \left(
                \sum_{m=0}^{M}
                \left( \sum_{\ell=0}^{M} C_{m,\ell}^{q} \right)^{\frac p q}
            \right)
            \| \nabla u \|_{L^{p}(\Omega)}^{p} 
            ,
        \end{split}
    \end{align*}
    where $q \in [1,\infty]$ satisfies $1 = 1/p + 1/q$ and with obvious modifications if $p=1$ or $p=\infty$. 
\end{proof}
